---
title: "Asthma project"
output: html_document
---
## Overall Goal
Identify disease-relevant cell types and genes for Asthma
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
source("code/make_plots.R")
```
## Enrichment estimates for individual annotation
### Adult-onset asthma
* GWAS: *Zhu et al. 2019* (case N=22296, control N=347481)  
* cell-type specific Annotations: *Zhang et al.2021* - single-cell ATAC-Seq data for adult lungs
    + Entropy based method used to identify cell-type specific peaks
    + For each cell type, peaks that pass 0.1% FDR cutoff were used to generate binary annotations

### Cell compositions for adult lung tissues
```{r echo=F}
cell_dict<-read.table("output/asthma/lung_clusters_dict.txt", header=T)
#Distributions of number of cell-type specific peaks across tissues
lungs<-read.table("output/asthma/lung_clusters_info.txt", header=T)
lungs<-lungs[-1]
df<-lungs[lungs>5]
names(df)<-names(lungs)[which(lungs>5)]
df<-df[1:(length(df)-1)]
gginput<-data.frame(names(df), df)
colnames(gginput)<-c("term", "counts")
gginput["cell_type"]<-unlist(lapply(gginput$term, function(i){cell_dict[cell_dict$X1==i,]$X2}))
ggplot(gginput, aes(x=cell_type, y=counts))+geom_bar(stat="identity") + theme(axis.text.x=element_text(angle=90, vjust = +0.7))
```

### Enrichment results

```{r echo=F}
df<-read.table("output/asthma/celltype_specific_adult_lungs_torus.est",header = F)
colnames(df)<-c("term", "est", "low", "high")
df["cell_type_raw"]<-unlist(lapply(df$term, function(i){strsplit(i, "_atac")[[1]][1]}))
df['highlight']<-df$low >=0
df["cell_type"]<-unlist(lapply(df$cell_type_raw, function(i){cell_dict[cell_dict$X1==i,]$X2}))
plot_torus_enrichment(df, text_angle=90)
```
Log2 scaled enrichment estimates for each cell type, which was individually run using Torus. The highlighted bars show significant enrichment, while the bars in grey color do not. Asthma-associated variants are significantly enriched in certain sub-clusters of macrophages and fibroblasts, but not in others. The cell types that show the highest enrichment in the chromatin accessibility peaks from immune-related cells.

* *Caldero et al. 2019* - ATAC-Seq data for FACS-sorted cells from whole blood
    - Significant differentially accessible regions when compared to progenitor cells
    - Significant differentially accessible regions under stimulation

```{r echo=F}
df<-read.table("output/asthma/diffe_adult_blood_torus.est", header = F)
colnames(df)<-c("term", "est", "low", "high")
df['highlight']<-df$low >=0
df["cell_type"]<-unlist(lapply(df$term, function(i){strsplit(i, "_U")[[1]][1]}))
#omit cell types with very large standard deviations
df['omit']<-df$low < -10 | df$high > 10
plot_torus_enrichment(df[df['omit']==F,], text_angle=60, vjust=0) + ggtitle("Differentiation-specific & accessible regions in blood")
```

* Not sure why much fewer cell types show differentially accessiblity during differentiation. 

```{r echo=F}
df<-read.table("output/asthma/stimu_adult_blood_torus.est",header = F)
colnames(df)<-c("term", "est", "low", "high")
df["cell_type"]<-unlist(lapply(df$term, function(i){strsplit(i, "_S")[[1]][1]}))
df['highlight']<-df$low >=0
#omit cell types with very large standard deviations
df['omit']<-df$low < -10 | df$high > 10
plot_torus_enrichment(df[df['omit']==F,], text_angle=90, vjust=1) + ggtitle("Stimulation-specific & accessible regions in blood")

```
Upon stimulation, the enrichment signals prevail across a broader range of immune cells. There are more diverse stimulated cell types that show differential accessibility against the unstimulated ones. 

## Joint enrichment estimates across annotations 

### immune cells in lungs vs. immune cells in blood
**regions with differentiation-specific accessibility in blood**
```{r echo=F}
df<-read.table("output/asthma/joint_lung_vs_blood_immune_diff_torus.est",header = F)
colnames(df)<-c("term", "est", "low", "high")
df<-df[-1,]
df["cell_type_raw"]<-unlist(lapply(df$term, function(i){strsplit(i, "_atac")[[1]][1]}))
df["cell_type"]<-unlist(lapply(df$cell_type_raw, function(i){
  if(i %in% cell_dict$X1){
    return(cell_dict[cell_dict$X1==i,]$X2)
  }else{
    return(i)
  }
}))
df['highlight']<-df$low >=0
#omit cell types with very large standard deviations
df['omit']<-df$low < -15 | df$high > 10
plot_torus_enrichment(df[df['omit']==F,], text_angle=90, vjust=1) + ggtitle("Blood differentiation-specific & accessible regions")
```
Both clusters of cell-type specific accessible regions from T lymphocytes in lungs are significantly enriched with risk variants conditional on differentiation-specific regions identified from blood. 

```{r echo=F}
df<-read.table("output/asthma/joint_lung_vs_blood_immune_stimu_torus.est",header = F)
colnames(df)<-c("term", "est", "low", "high")
df<-df[-1,]
df["cell_type_raw"]<-unlist(lapply(df$term, function(i){strsplit(i, "_atac")[[1]][1]}))
df["cell_type"]<-unlist(lapply(df$cell_type_raw, function(i){
  if(i %in% cell_dict$X1){
    return(cell_dict[cell_dict$X1==i,]$X2)
  }else{
    return(i)
  }
}))
df['highlight']<-df$low >=0
#omit cell types with very large standard deviations
df['omit']<-df$low < -15 | df$high > 10
plot_torus_enrichment(df[df['omit']==F,], text_angle=90, vjust=1) + ggtitle("Stimulation-specific & accessible regions in blood")
```
The enrichment signal of T lymphocytes remains after conditional on stimulated immune cell types in blood. The additional enrichment signals come from stimulated memoryCD8+ T-cells, T_helpler cells, mast cells, Th1 precursors, and Th2 precursors. 

### Children-onset Asthma
* GWAS: *Zhu et al. 2019*  
* cell-type specific Annotations: *Domcke et al.2020* - single-cell ATAC-Seq data for fetal lungs
```{r echo=F}
# lung_cells<-c("Bronchiolar_alveolar_epithelial_cell",
#               "Ciliated_epithelial_cell",
#               "Lymphatic_endothelial_cell",
#               "Lymphoid_cell",
#               "Myeloid_cell",
#               "Neuroendocrine_cell",
#               "Megakaryocyte",
#               "Neuroendocrine_cell",
#               "Stromal_cell_lineage"
#               )
# df<-read.table(paste(workdir, "asthma_child/torus_enrichment_all.est", sep="/"),
#               header = F)
# colnames(df)<-c("term", "est", "low", "high")
# df["cell_type"]<-unlist(lapply(df$term, function(i){strsplit(i, "_atac")[[1]][1]}))
# df['highlight']<-df$low >=0
# plot_torus_enrichment(df[df$cell_type %in% lung_cells,], text_angle=45, vjust=0)
```